---
title: "VIS2128A45"
author: "DM"
date: "9/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Java and Libraries



```{r cars}
options(java.parameters = "-Xmx2G")

library(r5r)
library(osmextract)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
library(tigris)
library(wesanderson)
library(tidytransit)
library(stars)
```

## Download File

You can also embed plots, for example:

```{r}

dir.create("networks")

download.file("https://cdn.mbta.com/MBTA_GTFS.zip", file.path("networks","Bostongtfs.zip"), mode = "wb", quiet=TRUE)

```

## Visualize Street Network of Chelsea



```{r cars}
chelsea_file <- oe_match("Chelsea, Massachusetts")

chelsea_streets <- oe_read(chelsea_file$url, 
                   provider = "openstreetmap_fr", 
                   download_directory = "networks", 
                   layer = "lines", 
                   quiet = TRUE) %>%
  filter(!is.na(highway))

ggplot(chelsea_streets) + 
  geom_sf()
```

## Setting Chelsea City Boundaries



```{r}
chelsea_tigris <- places("Massachusetts")%>%
  filter(NAME == "Chelsea") %>%
  st_transform(crs = st_crs(chelsea_streets))

ggplot(chelsea_tigris) +
  geom_sf()
```

## Loading Better Chelsea Map



```{r cars}
chelsea <- st_read("tufts-chelseaborderpoly05-geojson.json")

ggplot(chelsea) +
  geom_sf()
```

## Loading Massachusetts Coordinate Plane and Chelsea Streets



```{r cars}
MA_state_plane <- "+proj=lcc +lat_1=42.68333333333333 +lat_2=41.71666666666667 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"

chelsea_streets <- chelsea_streets %>%
  st_transform(MA_state_plane)

chelsea <- chelsea %>%
  st_transform(MA_state_plane)

chelsea_streets <- chelsea_streets[chelsea,]

ggplot(chelsea_streets) +
  geom_sf()
```

## Adding Aesthetics to Chelsea Transportation Map



```{r cars}
ggplot() +
  geom_sf(data = chelsea, fill = "azure", color = "gray") + 
  geom_sf(data = chelsea_streets, alpha = 0.1) +
  theme_map()
```

### Adding Park Points


```{r cars}
chelsea_greenspaces <- oe_read(chelsea_file$url,
                  provider = "openstreetmap_fr",
                  download_directory = "networks",
                  layer = "points",
                  quiet = TRUE) %>%
  filter(str_detect(other_tags, '"leisure"=>"park"')) %>%
  st_transform(crs = MA_state_plane) %>%
  st_filter(chelsea) %>% 
  rename(id = osm_id)

chelsea_greenspaces <- st_read("tufts-chelseaparksply05-geojson.json")

greenspaces_point <- st_make_valid(chelsea_greenspaces) %>%
  st_centroid(chelsea_greenspaces)
  
ggplot() +
  geom_sf(data = chelsea, fill = "#758B66", color = NA, alpha = 0.8) + 
  geom_sf(data = greenspaces_point)+
  theme_map()
```

## Making a Grid



```{r cars}
grid <- st_sf(st_make_grid(chelsea, square = FALSE,
                           n = c(100, 100),
                           what = "polygons")) %>% 
  st_filter(chelsea)

colnames(grid) <- "geometry"
st_geometry(grid) <- "geometry"

grid <- grid %>% 
  mutate(id = seq(1, length(grid$geometry), by = 1))


ggplot() +
  geom_sf(data = grid)+
  theme_map()
```

## Centroids and Parks



```{r cars}
grid_points <- st_centroid(grid)

ggplot() +
  geom_sf(data = grid_points, size = 0.75) +
  geom_sf(data = greenspaces_point, color = "yellow1") +
  theme_map()
```


## Set Up R5R



```{r cars}
r5r_core <- setup_r5("networks", verbose = FALSE)
```


## Calculating Travel Time Matrix



```{r cars}
ttm <- travel_time_matrix(r5r_core = r5r_core,
                          origins = st_transform(x = greenspaces_point, crs = "WGS84"),
                          destinations = st_transform(x = grid_points, crs = "WGS84"),
                          mode = "WALK",
                          departure_datetime = as.POSIXct("15-11-2021 14:00:00",
                                 format = "%d-%m-%Y %H:%M:%S"),
                          max_walk_dist = 1000,
                          max_trip_duration = 480,
                          verbose = FALSE)

tt_wide <- ttm %>%
  pivot_wider(names_from = fromId, 
              names_prefix = "from", values_from = travel_time) %>%
  rename(id = toId) %>% 
  merge(grid) %>%
  replace(is.na(.), 999) %>%
  rowwise() %>%
  mutate(from_any = min(c_across(starts_with("from")), na.rm = TRUE))

st_geometry(tt_wide) <- "geometry"

ggplot() +
  geom_sf(data = tt_wide, 
          aes(fill = from_any), 
          color = NA) +
  geom_sf(data = greenspaces_point, alpha = 0.2) +
  scale_fill_gradientn(colors = wes_palette(name = "Zissou1", n= 3, type = "continuous"),
        name = "Walking time to\nthe nearest green space \nstructure (minutes)",
        position = "right") +
  coord_sf(crs = MA_state_plane) +
  theme_map()+
  theme(legend.position = "right")

```


## Creating Isochrones



```{r cars}
iso_pallete <- wes_palette("Zissou1", n = 5)

iso10min <- tt_wide[tt_wide$from_any < 11,] %>%
  st_union()

iso20min <- tt_wide[tt_wide$from_any < 21,] %>%
  st_union()

iso30min <- tt_wide[tt_wide$from_any < 31,] %>%
  st_union()

ggplot(chelsea_streets) +
  geom_sf(data = iso30min, 
          aes(fill = "Area within 30 minutes"), 
          color = NA) +
  geom_sf(data = iso20min, 
          aes(fill = "Area within 20 minutes"), 
          color = NA) +
  geom_sf(data = iso10min, 
          aes(fill = "Area within 10 minutes"), 
          color = NA) +
  geom_sf(alpha = 0.5) +
  scale_fill_manual(values = c(iso_pallete[1], 
                               iso_pallete[3],
                               iso_pallete[5]),
        name = "Transit Travel\ntime to the\nnearest green space\n(minutes)") +
  coord_sf(crs = MA_state_plane) +
  theme_map()
```



```{r}
ggplot(chelsea_transit_stops) +
  geom_sf()
```

```{r}
ggplot(chelsea_transit_grid) +
  geom_sf()
```


```{r}
ggplot(grid_points) +
  geom_sf()
```

## Calculating Accesibility



```{r cars}
chelsea_transit <- read_gtfs(file.path("networks","Bostongtfs.zip"))

chelsea_transit_stops <- st_as_sf(chelsea_transit$stops %>% 
                                    drop_na(c("stop_lat", "stop_lon")), 
                          coords = c("stop_lon", "stop_lat")) %>% 
                          st_transform(crs = st_crs(grid_points))

chelsea_transit_grid <- grid %>%
  mutate(num_stops = lengths(st_covers(grid, chelsea_transit_stops)))

chelsea_transit_points <- st_centroid(chelsea_transit_grid)

ggplot(chelsea_transit_points) +
  geom_sf(aes(color = as.character(num_stops))) +
  scale_color_manual(values = c("gray", "cornsilk", "lightgreen"), 
                    name = "Number of\ntransit stops")
```



```{r}
chelsea_transit_access <- accessibility(r5r_core,
                        origins = chelsea_transit_points,
                        destinations = chelsea_transit_points,
                        mode = "WALK",
                        opportunities_colname = "num_stops",
                        decay_function = "step",
                        cutoffs = 11,
                        departure_datetime = as.POSIXct("15-11-2021 14:00:00",
                                 format = "%d-%m-%Y %H:%M:%S"),
                        max_walk_dist = 500,
                        time_window = 60,
                        percentiles = 50,
                        verbose = FALSE) %>%
  mutate(id = as.numeric(from_id)) %>%
  merge(grid)

st_geometry(chelsea_transit_access) <- "geometry"

ggplot(chelsea_transit_access) +
  geom_sf(aes(fill = accessibility), color = NA) +
  scale_fill_viridis_c(name = "Transit stops\nwithin 10-minutes\nwalk") +
  coord_sf(crs = MA_state_plane) +
  theme_void()
```


```{r}
chelsea_transit_access2 <- accessibility(r5r_core,
                        origins = chelsea_transit_points,
                        destinations = chelsea_transit_points,
                        mode = "WALK",
                        opportunities_colname = "num_stops",
                        decay_function = "exponential",
                        cutoffs = 5,
                        departure_datetime = as.POSIXct("15-11-2021 14:00:00",
                                 format = "%d-%m-%Y %H:%M:%S"),
                        max_walk_dist = 500,
                        time_window = 60,
                        percentiles = 50,
                        verbose = FALSE) %>%
  mutate(id = as.numeric(from_id)) %>%
  merge(grid)

st_geometry(chelsea_transit_access2) <- "geometry"

ggplot(chelsea_transit_access2) +
  geom_sf(aes(fill = accessibility), color = NA) +
  scale_fill_viridis_c(name = "Accessiblity score") +
  coord_sf(crs = MA_state_plane) +
  theme_void()
```


```{r}
stop_r5(r5r_core)
```














